---
title: "Ren V _Final_Iris Project"
author: "Sattiki"
date: "2024-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load renv and initialize it: Here, we are setting up the environment
```{r}
install.packages("renv")  
library(renv)
renv::init() #The renv::init() function, creates a lock in the project directory so only the relevant packages to the directory will be accessed.
```



# Install the packages needed for the project
```{r}
install.packages("dplyr")
install.packages("ggplot2")
install.packages("summarytools")

```
# Snapshot to save the environment in renv.lock
```{r}
renv::snapshot() #The renv::snapshot() function will capture the package versions and store them in the renv.lock file, ensuring that the exact versions can be reproduced.
```

```{r}
library(dplyr)
library(ggplot2)
library(summarytools)
```



# Load the iris dataset
```{r}
data("iris")
```

# Show basic information and structure

```{r}
summary(iris)
dfSummary(iris)  # Provides a detailed summary using summarytools

```

# Calculate the average sepal and petal lengths by species
```{r}
iris_summary <- iris %>%
  group_by(Species) %>%
  summarise(
    Avg_Sepal_Length = mean(Sepal.Length),
    Avg_Sepal_Width = mean(Sepal.Width),
    Avg_Petal_Length = mean(Petal.Length),
    Avg_Petal_Width = mean(Petal.Width)
  )

print(iris_summary)
```



5. Data Visualization

```{r}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point() +
  labs(title = "Sepal Length vs Petal Length",
       x = "Sepal Length (cm)",
       y = "Petal Length (cm)") +
  theme_minimal()
```

#Restoring the Environment (For Collaboration): If we share this project with someone else, they can recreate the exact environment by running:


# Restore the project environment based on renv.lock


```{r}
renv::restore() #This command installs all necessary packages in the specific versions used for the project, allowing others to reproduce your results exactly.
```

